Performance comparison of Dask and Apache Spark on HPC systems

### 2022-02-23 latex/paper.tex
Minor fixes

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences
and similarities\HL{, for HPC,} between Dask and Spark for data-intensive applications.

@@

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a switch.
	The login node, hosting an NFS, is also connected by a {\SI{10}{\giga\bit\per\second}} Ethernet link to the compute nodes.}

@@

\HL{Our system uses similar hardware as Compute Canada Beluga, and replicates as
	close as possible it's configuration.
	This let us make a fair comparison of Spark and Dask in the condition of a realistic
HPC system.}

@@

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel
intensities of the BigBrain image. We set the number of clusters to 3. The
application starts by reading the image blocks, combining all voxels in a
1-D array, and choosing initial centroids using the min, max, and
intermediate values. It assigns each voxel to its closest centroid and
updates each centroid with the average of its assigned voxels. It repeats
the assignment and update steps for a configurable number of iterations.
Finally, the voxels of the image blocks are classified and written back to
the file system. Updating the centroids involves substantial data
communication between the workers.
The computation to find the closest centroids for each voxels scales with the number of centroids.
It can be computed in parallel at the cost of higher memory or sequentially to reduce memory usage.
We decided to compute the nearest centroids in parallel since we use only three centroids.

\HL{Due to resources limitation on our infrastructure, the Kmeans experiments
	are performed with only a sample of half the BigBrain dataset ({\SI{303}{\gibi\byte}}).}

@@

From Figure \ref{fig:increment_worker}, we also observe that Dask's overhead is higher than Spark's.
While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
\HL{Although Dask has a higher overhead, the total time for both engines is overall significantly similar.}
This is due to a tradeoff between disk bandwidth and compute/overhead time.
On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
\HL{The only significant difference in total time is when two nodes are used.
	This results in an approximately 3\% total runtime difference.}
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
	\caption{Increment: varying nodes -- 5000 files, 5 iterations}
	\label{fig:increment_worker}
\end{figure}

@@

From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
However, the total execution time of the application is \HL{significantly equivalent} due to the tradeoff between I/O and compute/overhead time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_block.jpg}
	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations}
	\label{fig:increment_block}
\end{figure}

@@

\section{Discussion}
\subsection{I/O bottleneck}
\label{subsec:io-bottleneck}
We used Lustre as shared file system as it is a \HL{de facto} solution in HPC systems.
The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
\HL{In this paper, the performance scale when increasing the number of nodes, although not linear due to an I/O bottleneck.
	The system was designed such that the filesystem has similar speed to the network.
Upgrading to more performant networking and storage compononents would reduce the bottleneck seen when increasing the number of workers.}

>>>>

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences
and similarities between Dask and Spark for data-intensive applications \HL{executed on HPC systems}.

@@

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a Mellanox MSN2410 switch.
	The login node, hosting an NFS server, is connected to the compute nodes through the same switch, with {\SI{10}{\giga\bit\per\second}} Ethernet.}

@@

\HL{Our system uses similar hardware as Compute Canada's B\'eluga cluster --- a representative cluster available to researchers in Canada --- and replicates as
	close as possible its configuration.
	This let us compare Spark and Dask in the condition of a realistic
HPC system.}

@@

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel
intensities of the BigBrain image. We set the number of clusters to 3. The
application starts by reading the image blocks, combining all voxels in a
1-D array, and choosing initial centroids using the min, max, and
intermediate values. It assigns each voxel to its closest centroid and
updates each centroid with the average of its assigned voxels. It repeats
the assignment and update steps for a configurable number of iterations.
Finally, the voxels of the image blocks are classified and written back to
the file system. Updating the centroids involves substantial data
communication between the workers.
The computation to find the closest centroids for each voxels scales with the number of centroids.
It can be computed in parallel at the cost of higher memory or sequentially to reduce memory usage.
We decided to compute the nearest centroids in parallel since we use only three centroids.
\HL{Due to resources limitation on our infrastructure, the Kmeans experiments
	are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).}

@@

From Figure \ref{fig:increment_worker}, we also observe that Dask's overhead is higher than Spark's.
While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
\HL{Although Dask has a higher overhead, the total time for both engines is overall similar.}
This is due to a tradeoff between disk bandwidth and compute/overhead time.
On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
\HL{The only significant difference in total time is when two nodes are used.
	This results in an approximately 3\% total runtime difference.}
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
	\caption{Increment: varying nodes -- 5000 files, 5 iterations}
	\label{fig:increment_worker}
\end{figure}

@@

From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
However, the total execution time of the application is \HL{equivalent} due to the tradeoff between I/O and compute/overhead time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_block.jpg}
	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations}
	\label{fig:increment_block}
\end{figure}

@@

\section{Discussion}
\subsection{I/O bottleneck}
\label{subsec:io-bottleneck}
We used Lustre as shared file system as it is a \HL{de facto} solution in HPC systems.
The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
\HL{In this paper, the performance scales when increasing the number of nodes, although not linearly due to an I/O bottleneck.
	The system was designed such that the filesystem has similar speed to the network.
Upgrading to more performant networking and storage compononents would reduce the bottleneck seen when increasing the number of workers.}
